---
title: "Untitled"
author : "Jeffrey Liang"
date : "10/02/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rnoaa)
library(ggridges)
library(patchwork) # use to multi plot, plt1 + (pl2 + plt3) / plt4
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d 
```


# Problem 1
```{r}
library(p8105.datasets)
data("instacart")

```


# Problem 2
## Load data

```{r}
accelerometer =
  read_csv(here::here("data/accel_data.csv")) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    cols = starts_with("activity_"),
    names_to = 'min',
    values_to = "activity",
    names_prefix = "activity_"
  ) %>% 
  mutate_at(c("min"),as.numeric) %>% 
  mutate(weekday_vs_weekend = day %in% c("Sunday","Saturday"),
         weekday_vs_weekend=
           case_when(weekday_vs_weekend ~"weekend",
                     !weekday_vs_weekend ~ "weekday") %>% 
           as.factor())
str(accelerometer)
skimr::skim_without_charts(accelerometer)
```
\ The original data is a "wider" format data with `r paste(dim(read_csv(here::here("data/accel_data.csv"))),collapse = " x ")` dimension. In order to make the data compatible with machine, pivot_long() is used to make the _activity_*_ of all subjects into columns of _min and activity_. Following instruction, weekend vs weekday's variable is build on _day_id_ and result into a `r paste(dim(accelerometer),collapse = " x ")` dataset.




